\huge \textbf{Words That Matter:\\ Vocabulary Pruning in ModernBERT}

### 2025-05-11 paper/main.tex
update gitignore

\begin{table*}[h]
\centering
\scriptsize
\caption{Comparison of vocabulary pruning techniques on GLUE benchmark tasks. Results show accuracy on the development set with varying pruning methods. Best results for each task are highlighted in \textbf{bold}.}
\label{tab:results}
\setlength{\tabcolsep}{2.8pt}
\begin{tabular}{lcccc|cc|cc|cc|cc|c}
\toprule
\textbf{Method} & \textbf{SST-2} & \textbf{MRPC} & \textbf{CoLA} & \textbf{STS-B} & \multicolumn{2}{c|}{\textbf{MNLI}} & \multicolumn{2}{c|}{\textbf{QQP}} & \multicolumn{2}{c|}{\textbf{QNLI}} & \multicolumn{2}{c|}{\textbf{RTE}} & \textbf{AVG} \\
\midrule
\multicolumn{14}{l}{\textit{Baseline}} \\
ModernBERT (Full) & 0.951 & 0.856 & 0.586 & 0.917 & \multicolumn{2}{c|}{0.881} & \multicolumn{2}{c|}{0.917} & \multicolumn{2}{c|}{0.939} & \multicolumn{2}{c|}{0.598} & 0.831 \\
\midrule
\multicolumn{14}{l}{\textit{Vocabulary Pruning}} \\
Parameter Reduction & 18.85\% & 18.56\% & 22.61\% & 18.76\% & \multicolumn{2}{c|}{6.74\%} & \multicolumn{2}{c|}{4.79\%} & \multicolumn{2}{c|}{6.42\%} & \multicolumn{2}{c|}{17.06\%} & \\
Train Tokens Only & 0.950 & 0.831 & 0.509 & 0.917 & \multicolumn{2}{c|}{0.883} & \multicolumn{2}{c|}{0.917} & \multicolumn{2}{c|}{0.915} & \multicolumn{2}{c|}{0.598} & 0.815 \\
\midrule
Parameter Reduction & 20.25\% & 20.02\% & 23.27\% & 20.18\% & 10.57\% & 20.02\% & 9.01\% & 20.02\% & 10.31\% & 20.02\% & 18.83\% & 20.02\% & \\
Random Selection & 0.911 & 0.798 & 0.470 & 0.845 & 0.832 & 0.504 & 0.902 & 0.780 & 0.895 & 0.669 & 0.522 & 0.566 & 0.772 \\
Clustering-Based & 0.899 & 0.714 & 0.000 & 0.786 & 0.712 & 0.501 & 0.875 & 0.774 & 0.836 & 0.510 & 0.510 & 0.566 & 0.667 \\
Frequency-Based & 0.943 & 0.812 & 0.467 & \textbf{0.905} & 0.880 & 0.858 & \textbf{0.916} & \textbf{0.904} & \textbf{0.920} & \textbf{0.902} & 0.542 & 0.546 & 0.798 \\
Attention-Based & \textbf{0.947} & \textbf{0.864} & 0.589 & 0.885 & 0.864 & - & 0.912 & - & 0.912 & - & 0.550 & - & 0.803 \\
TF-IDF Based & 0.933 & 0.807 & 0.520 & 0.901 & 0.875 & 0.860 & 0.900 & 0.898 & 0.917 & 0.909 & 0.606 & \textbf{0.574} & 0.807 \\
\\ [-6pt]
\cdashline{1-14}
\\[-6pt]
Freq + OOV & 0.938 & 0.828 & 0.540 & 0.906 & \textbf{0.881} & - & \textbf{0.916} & - & 0.918 & - & 0.538 & - & 0.808 \\
TF-IDF + OOV & 0.923 & 0.834 & \textbf{0.615} & 0.903 & 0.878 & - & 0.906 & - & 0.919 & - & 0.554 & - & \textbf{0.817} \\
\\ [-8pt]
\cdashline{1-14}
\\[-6pt]
\multicolumn{14}{l}{\textit{Encoder Layer Pruning}} \\
Parameter Reduction & 20.16\% & 20.16\% & 20.16\% & 20.16\% & \multicolumn{2}{c|}{20.16\%} & \multicolumn{2}{c|}{20.16\%} & \multicolumn{2}{c|}{20.16\%} & \multicolumn{2}{c|}{20.16\%} & \\
LoSparse & 0.929 & 0.856 & 0.316 & 0.882 & \multicolumn{2}{c|}{0.858} & \multicolumn{2}{c|}{0.911} & \multicolumn{2}{c|}{0.907} & \multicolumn{2}{c|}{\textbf{0.610}} & 0.784 \\
\bottomrule
\end{tabular}
\end{table*}

>>>>

\begin{table*}[h]
\centering
\scriptsize
\caption{Comparison of vocabulary pruning techniques on GLUE benchmark tasks. Results show accuracy on the development set with varying pruning methods. Best results for each task are highlighted in \textbf{bold}.}
\label{tab:results}
\setlength{\tabcolsep}{2.8pt}
\begin{tabular}{lcccc|cc|cc|cc|cc|c}
\toprule
\textbf{Method} & \textbf{SST-2} & \textbf{MRPC} & \textbf{CoLA} & \textbf{STS-B} & \multicolumn{2}{c|}{\textbf{MNLI}} & \multicolumn{2}{c|}{\textbf{QQP}} & \multicolumn{2}{c|}{\textbf{QNLI}} & \multicolumn{2}{c|}{\textbf{RTE}} & \textbf{AVG} \\
\midrule
\multicolumn{14}{l}{\textit{Baseline}} \\
ModernBERT (Full) & 0.951 & 0.856 & 0.586 & 0.917 & \multicolumn{2}{c|}{0.881} & \multicolumn{2}{c|}{0.917} & \multicolumn{2}{c|}{0.939} & \multicolumn{2}{c|}{0.598} & 0.831 \\
\midrule
\multicolumn{14}{l}{\textit{Vocabulary Pruning}} \\
Parameter Reduction & 18.85\% & 18.56\% & 22.61\% & 18.76\% & \multicolumn{2}{c|}{6.74\%} & \multicolumn{2}{c|}{4.79\%} & \multicolumn{2}{c|}{6.42\%} & \multicolumn{2}{c|}{17.06\%} & \\
Train Tokens Only & 0.950 & 0.831 & 0.509 & 0.917 & \multicolumn{2}{c|}{0.883} & \multicolumn{2}{c|}{0.917} & \multicolumn{2}{c|}{0.915} & \multicolumn{2}{c|}{0.598} & 0.815 \\
\midrule
Parameter Reduction & 20.25\% & 20.02\% & 23.27\% & 20.18\% & 10.57\% & 20.02\% & 9.01\% & 20.02\% & 10.31\% & 20.02\% & 18.83\% & 20.02\% & - \\
Random Selection & 0.911 & 0.798 & 0.470 & 0.845 & 0.832 & 0.504 & 0.902 & 0.780 & 0.895 & 0.669 & 0.522 & 0.566 & - \\
Clustering-Based & 0.899 & 0.714 & 0.000 & 0.786 & 0.712 & 0.501 & 0.875 & 0.774 & 0.836 & 0.510 & 0.510 & 0.566 & - \\
Frequency-Based & 0.943 & 0.812 & 0.467 & \textbf{0.905} & 0.880 & 0.858 & \textbf{0.916} & \textbf{0.904} & \textbf{0.920} & \textbf{0.902} & 0.542 & 0.546 & - \\
Attention-Based & \textbf{0.947} & \textbf{0.864} & 0.589 & 0.885 & 0.864 & - & 0.912 & - & 0.912 & - & 0.550 & - & - \\
TF-IDF Based & 0.933 & 0.807 & 0.520 & 0.901 & 0.875 & 0.860 & 0.900 & 0.898 & 0.917 & 0.909 & 0.606 & \textbf{0.574} & - \\
\\ [-6pt]
\cdashline{1-14}
\\[-6pt]
Freq + OOV & 0.938 & 0.828 & 0.540 & 0.906 & \textbf{0.881} & - & \textbf{0.916} & - & 0.918 & - & 0.538 & - & - \\
TF-IDF + OOV & 0.923 & 0.834 & \textbf{0.615} & 0.903 & 0.878 & - & 0.906 & - & 0.919 & - & 0.554 & - & - \\
\\ [-8pt]
\cdashline{1-14}
\\[-6pt]
\multicolumn{14}{l}{\textit{Encoder Layer Pruning}} \\
Parameter Reduction & 20.16\% & 20.16\% & 20.16\% & 20.16\% & \multicolumn{2}{c|}{20.16\%} & \multicolumn{2}{c|}{20.16\%} & \multicolumn{2}{c|}{20.16\%} & \multicolumn{2}{c|}{20.16\%} & 20.16\% \\
LoSparse & 0.929 & 0.856 & 0.316 & 0.882 & \multicolumn{2}{c|}{0.858} & \multicolumn{2}{c|}{0.911} & \multicolumn{2}{c|}{0.907} & \multicolumn{2}{c|}{\textbf{0.610}} & 0.784 \\
\bottomrule
\end{tabular}
\end{table*}
